---
title: "Introduction to the rfars package"
author: "Cesar R. Urteaga-Reyesvera"
date: "2017-08-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the rfars package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Setup, include = FALSE}
knitr::opts_chunk$set(# Collapse output blocks
                      collapse = TRUE, 
                      comment = "#>", 
                      fig.width = 7, 
                      fig.height = 7, 
                      fig.align = "center")
# We set the working directory where the compressed files are stored within the
# package.
WD <- system.file("extdata", package = "rfars")
knitr::opts_knit$set(root.dir = WD)
library(rfars)
```

This vignette describes what is the `rfars` package and how should be used.

## Description

The `rfars` package bundles a set of functions to analyze the data from the US National Highway Traffic Safety Administration's [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

## Usage

### Getting the data into a more manageable format

The function `fars_read` allows you to extract the data from a compressed format (`.bz2`) into a tibble object.  

```{r}
# Extracts the data from a compressed file.
library(rfars)
data_2013 <- fars_read("accident_2013.csv.bz2")
head(data_2013)
```

### Summarizing the car fatalities

In order to get the total number of fatalities by month, use the `fars_summarize_years` function.

```{r}
# Creates a table by month & year with the total number of fatalities in 2013 y 2014.
fars_summarize_years(c(2013, 2014))
```

### Visualizing the car fatalities by state

The function `fars_map_state` creates a map with the car fatalities in a US state and a specific year.

```{r}
# Plots the car fatalities for the state of Florida (12) in 2015. 
fars_map_state(12, 2015)
```
