---
title: "Introduction to the rfars package"
author: "Cesar R. Urteaga-Reyesvera"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the rfars package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(# Collapse output blocks
                      collapse = TRUE, 
                      comment = "#>", 
                      fig.width = 7, 
                      fig.height = 7, 
                      fig.align = "center")
library(rfars)
```

This vignettes documents what is the `rfars` package and how should be used.

## Description

The `rfars` package bundle a set of functions to analyze the the data from the US National Highway Traffic Safety Administration's [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

## Usage

### Getting the data into a more manageable format

The function `fars_read` allows to extract the data, which is in a compress format (i.e., `.bz2`), into a tibble object.  

```{r ExtractFarsData}
# Extract the data from the compressed file.
library(rfars)
data_2013 <- fars_read("accident_2013.csv.bz2")
head(data_2013)
```

### Summarizing the car fatalities

In order to get the number of fatalities by month use the `fars_summarize_years`.

```{r}
# Creates a table by month & year with the number of fatalities in 2013 y 2014.
fars_summarize_years(c(2013, 2014))
```

### Visualizing the car fatalities.

The function `fars_map_state` creates a map with the car fatalities in a US state and a specific year.

```{r}
# Plots the car fatalities for the state of Florida (12) in 2015. 
fars_map_state(12, 2015)
```
